---
title: "Homework 4 - SDS315"
author: "Alaina Gomez (agg3395)"
date: "2024-02-13"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
---

```{r, echo = FALSE, message = FALSE}
ebay <- read.csv("/Users/alainagomez/Desktop/ebay.csv")
nbc_pilots <- read.csv("/Users/alainagomez/Desktop/nbc_pilotsurvey.csv")
sclass <- read.csv("/Users/alainagomez/Desktop/sclass2011.csv")
gas_prices <- read.csv("/Users/alainagomez/Desktop/gasprices.csv")
library(dplyr)
library(ggplot2)
library(tidyverse)
library(mosaic)
```

\newpage

[**Link to Github**](https://github.com/AlainaGomez/Homework-4---SDS315)

# Problem 1

## A

**Claim**: Gas stations charge more if they lack direct competition in sight (no close by gas competitors).

```{r, echo = FALSE, include = FALSE}
boot_comp <- do(10000)*diffmean(Price ~ Competitors, data=mosaic::resample(gas_prices))
boot_c <- boot_comp$diffmean
```

**Evidence**: After bootstrapping the sample 10,000 times, I was able to find the average gas price difference in gas stations with competitors and those without. The histogram compares the difference of average gas prices for competitive and non-competitive gas stations for 10,000 samples. With 95% confidence, the true difference between gas prices of these gas stations is between `r round(quantile(boot_c, probs = 0.025), 3)` and `r round(quantile(boot_c, probs = 0.975),3)` cents. Therefore, since 0 is in the confidence interval, there is not a significant difference between gas station gas prices if they do or do not have direct competition in sight.

```{r, echo = FALSE, message = FALSE}
ggplot(boot_comp) + geom_histogram(aes(x=diffmean), col = "black") + labs(x = "Average Differences in Price", y = "Count", title = "Competitive vs. Non-Competitive Gas Station Price Differences")
```

**Conclusion**: While there might be a slight average difference close to `r round(mean(boot_comp$diffmean),3)`, there is not a significant one. The histogram supports this idea as it is only slightly off from 0. Gas stations do NOT charge more if they lack direct competition in sight.

\newpage

## B

**Claim**: The richer the area (higher income), the higher the gas prices.

```{r, echo = FALSE}
boot_rich <- do(10000)*lm(Price ~ Income, data=mosaic::resample(gas_prices))
confint(boot_rich, level = 0.95)
```

**Evidence**: Boostrapping another sample 10,0000 times, I found the linear slope between local median income and gas prices of certain ZIP codes in Austin. After creating a confident interval from this relationship, I found that there is an increase between 0.00000006801 and 0.0000017729 cents per additional dollar increase of income. However, it can be said with 95% confidence that 0 is not within this confidence interval. As well as, when you create a jitter plot based on the original household median incomes and local gas prices, it is very apparent that there is little relationship between these two variables.

```{r, echo = FALSE, message = FALSE}
ggplot(gas_prices) + geom_jitter(aes(x = Income, y = Price), alpha=0.1) + 
  geom_smooth(aes(x=Income, y=Price), method='lm') + labs(x = "Income", y = "Price", title = "Household Median Income and Local Gas Prices")
```

**Conclusion**: Although 0 is in the confidence interval and there is technically a statistical difference between the variables, there is not a practical one as the difference is incredibly small to the average person. Therefore, richer areas (higher household median incomes) do NOT mean that there are higher gas prices.

\newpage

## C

**Claim**: Gas stations at stoplights charge more.

```{r, echo = FALSE}
boot_light <- do(10000)*diffmean(Price ~ Stoplight, data=mosaic::resample(gas_prices))
boot_l <- boot_light$diffmean
```

**Evidence**: I created another 10,000 sample bootstrap that compared the average gas price difference in gas stations that are located at/by stoplights and those that are not. The histogram compares the difference of average gas prices for stoplight and non-stoplight gas stations for 10,000 samples. With 95% confidence, the true difference between gas prices of these gas stations is between `r round(quantile(boot_l, probs = 0.025), 3)` and `r round(quantile(boot_l, probs = 0.975),3)` cents. Because 0 cents is within this confidence interval, there is not a significant difference between gas station gas prices if they are at a stoplight or not.

```{r echo=FALSE, message=FALSE}
ggplot(boot_light) + geom_histogram(aes(x=diffmean), col = "black") + labs(x = "Average Differences in Price", y = "Count", title = "Stoplight vs. Non-Stoplight Gas Station Price Differences")
```

**Conclusion**: The histogram clearly shows a large portion of the differences are close to 0; meaning no real difference in price. Gas stations at stoplights do NOT charge more than those that are not at gas stations.

\newpage

## D

**Claim**: Gas stations with direct highway access charge more.

```{r, echo = FALSE}
boot_high <- do(10000)*diffmean(Price ~ Highway, data=mosaic::resample(gas_prices))
boot_h <- boot_high$diffmean
```

**Evidence**: Bootstrapping another sample 10,000 times, I was able to find the average gas price difference in gas stations with highway access and those without. The histogram compares the difference of average gas prices for highway access and non-highway access gas stations for 10,000 samples. With 95% confidence, the true difference between gas prices of these gas stations is between `r round(quantile(boot_h, probs = 0.025), 3)` and `r round(quantile(boot_h, probs = 0.975),3)` cents. Because 0 IS NOT within this confidence interval, this signals that there is a statistically significant difference in gas prices for gas stations that have highway access in comparison to those that do not.

```{r, echo = FALSE, message = FALSE}
ggplot(boot_high) + geom_histogram(aes(x=diffmean), col = "black") + labs(x = "Average Differences in Price", y = "Count", title = "Highway vs. Non-Highway Gas Station Price Differences")
```

**Conclusion**: Due to the significance of average differences in price, the histogram supports the claim that gas stations with direct highway access charge slightly more with an average of about `r round(mean(boot_high$diffmean),3)` cents more. Gas stations with direct highway access DO technically charge more than their counterparts that do not have highway access. However, this slight difference may not be that important to day-to-day people as about 4 cents may not matter to the average person.

\newpage

## E

**Claim**: Shell charges more than all other non-Shell brands.

```{r, echo = FALSE}
gas_prices <- mutate(gas_prices, is_shell = ifelse(Brand == "Shell", "Yes", "No"))
boot_shell <- do(10000)*diffmean(Price ~ is_shell, data=mosaic::resample(gas_prices))
boot_s <- boot_shell$diffmean
```

**Evidence**: After bootstrapping 10,000 samples, I found the average gas price difference of Shell gas stations and non-shell gas stations. The histogram compares the difference between average gas prices for gas stations owned by Shell and those that are not. With 95% confidence, the true difference of these gas stations is between `r round(quantile(boot_s, probs = 0.025),3)` and `r round(quantile(boot_s, probs = 0.975),3)` cents. Because 0 is in this confidence interval, there is no significance difference in gas price depending on whether or not a gas station is Shell or not.

```{r, echo = FALSE, message = FALSE}
ggplot(boot_shell) + geom_histogram(aes(x=diffmean), col = "black") + labs(x = "Average Differences in Price", y = "Count", title = "Shell vs. Non-Shell Gas Station Price Differences")
```

**Conclusion**: While the histogram is slightly off with an average price difference of about `r round(mean(boot_shell$diffmean),3)` cents, there is not a significant difference. Shell does NOT charge more for gas than all other non-Shell gas stations.

\newpage

# Problem 2

## A

```{r, echo = FALSE}
part_a <- filter(sclass, year == 2011)
part_a <- filter(sclass, trim == "63 AMG")
boot_ax <- do(10000)*mean(~mileage, data=mosaic::resample(sclass))
```

```{r, echo = FALSE, message = FALSE}
ggplot(boot_ax) + geom_histogram(aes(x=mean), col = "black") + labs(x = "Average Mileage", y = "Count", title = "Average mileage of 2011 S-Class
63 AMGs")
options(scipen = 999)
```

With 95% confidence, the true average mileage of the 2011 S-Class 63 AMGs that were hitting the used-car market at this time is between `r round(quantile(boot_ax$mean, probs = 0.025), 2)` and `r round(quantile(boot_ax$mean, probs = 0.975), 2)`.

\newpage

## B

```{r, echo = FALSE}
part_b <- filter(sclass, year == 2014)
part_b <- filter(sclass, trim == "550")
sclass <- mutate(sclass, is_black = ifelse(color == "Black", "Yes", "No"))
boot_bx <- do(10000)*prop(~is_black, data=mosaic::resample(sclass))
boot_b <- boot_bx$prop_No
```

```{r, echo = FALSE, message = FALSE}
ggplot(boot_bx) + geom_histogram(aes(x=prop_No), col = "black") + labs(x = "Proportion", y = "Count", title = "Proportion of all 2014 S-Class 550s painted Black")
```

With 95% confidence, the true the proportion of all 2014 S-Class 550s that were painted black is between `r round(quantile(boot_b, probs = 0.025),3)` and `r round(quantile(boot_b, probs = 0.975),3)`

\newpage

# Problem 3

## A

```{r, echo = FALSE}
ed_and_earl <- filter(nbc_pilots, Show == "Living with Ed" | Show == "My Name is Earl")
happy_mean <- do(10000)*diffmean(Q1_Happy ~ Show, data=mosaic::resample(ed_and_earl))
```

**Question**: Is there evidence that one show consistently produces a higher average Q1_Happy response among viewers?

**Approach**: I used the bootstrapping method to resample the differences of average happiness ratings of "Living with Ed" and "My Name is Earl" 10,000 times. Resampling this many times allowed me to find the true response differences between the show among viewers.

**Results**: The histogram compares the average difference of happiness ratings from the resampling. While there is a slight difference, 0 is within the 95% confidence interval of the data. This fact is also shown in the function that computes confidence intervals based on the Central Limit Theorem. The average difference of the histogram is about `r round(mean(happy_mean$diffmean),3)`.

```{r, echo = FALSE, message = FALSE}
ggplot(happy_mean) + geom_histogram(aes(x = diffmean), col = "black") + labs(x = "Q1_Happy Average Show Differences", y = "Count", title = "Q1_Happy Average Show Differences")
confint(happy_mean$diffmean, level = 0.95)
```

**Conclusion**: Because 0 is in the confidence interval, there is not a statistically significant difference between Q1_Happy responses from the shows. Therefore, it can be said with 95% confidence that there is no perceived happiness difference between the "Pilot" episodes of these shows.

\newpage

## B

```{r, echo = FALSE}
loser_and_la <- filter(nbc_pilots, Show == "The Apprentice: Los Angeles" | Show == "The Biggest Loser")
annoyed_mean <- do(10000)*diffmean(Q1_Annoyed ~ Show, data=mosaic::resample(loser_and_la))
```

**Question**: Is there evidence that one show consistently produces a higher mean Q1_Annoyed response among viewers?

**Approach**: I used another bootstrap to resample the differences of average annoyed ratings of "The Apprenctice: Los Angeles" and "The Biggest Loser" 10,000 times. Resampling this many times allowed me to find the true response differences between the show among viewers.

**Results**: The histogram compares the average difference of annoyed ratings from the resampling. It is clear from the auto-generated "large-sample" confidence intervals that 0 is not within the 95% confidence interval. As well as, 0 is not within 95% confidence of the histogram. The average difference of the histogram is about `r round(mean(annoyed_mean$diffmean),3)`.

```{r, echo = FALSE, message = FALSE}
ggplot(annoyed_mean) + geom_histogram(aes(x = diffmean), col = "black") + labs(x = "Average Annoyed Rating Difference", y = "Count", title = "Q1_Annoyed Average Show Differences")
confint(annoyed_mean$diffmean, level = 0.95) 
```

**Conclusion**: Because 0 is not in the confidence interval, there is evidence that one show consistently produces a higher Q1_Annoyed response than the other amongst viewers. It is clear that one show's "Pilot" episode annoyed more viewers than the other did to the average viewer.

\newpage

## C

```{r, echo = FALSE}
dance_stars <- filter(nbc_pilots, Show == "Dancing with the Stars")
# those with "Yes" for confusing rating have a rating less than 4 on Q2_Confusing
dance_stars <- mutate(dance_stars, confusing_rating = ifelse(Q2_Confusing < 4, "Yes", "No"))
dance_prop <- do(10000)*prop(~confusing_rating, data = mosaic::resample(dance_stars))
```

**Question**: What proportion of American TV watchers would we expect to give a response of 4 or greater to the "Q2_Confusing" question?

**Approach**: I took the approach to bootstrap 10,000 samples of the Dancing with the Stars confusing rating, Firs, I created a new variable that decided if the Q2_Confusing question had a rating greater than or equal to 4. Then, I bootstrapped according to the proportion of viewers that found the show confusing (Rating \>= 4) and those that did not.

**Results**: The histogram shows the distribution of proportions that viewers found of the Dancing with the Stars show based on the Q2_Confusing question. It is clear that in both the histogram and the generated confidence interval that the average proportion of American TV Watchers to give the show a confusing ratings greater than or equal to four is around `r round(mean(dance_prop$prop),3)`.

```{r, echo = FALSE}
ggplot(dance_prop)  + geom_histogram(aes(x = prop_No), col = "black", bins = 40) + labs(x = "Proportion", y = "Count", title = "Proportion of Q2_Confusing Ratings >= 4")
# hide messages from showing
suppressMessages(confint(dance_prop$prop_No, level = 0.95)) 
```

**Conclusion**: It can be said with 95% confidence that the proportion of viewers that rate Dancing with the Star's Q2_Confusing question with a rating greater than or equal to four is between `r round(quantile(dance_prop$prop, probs = 0.025),3)` and `r round(quantile(dance_prop$prop, probs = 0.975),3)`. This rating is equivalent to an average of about `r 100*(round(mean(dance_prop$prop),2))`% of viewers finding the show confusing.

\newpage

# Problem 4

```{r, echo = FALSE}
boot_ebay <- do(10000)*diffmean(rev_ratio ~ adwords_pause, data = mosaic::resample(ebay))
```

**Question**: Is the revenue ratio is the same in the treatment and control groups? Or if the data favors the idea that paid search advertising on Google creates extra revenue for EBay.

**Approach**: In order to answer this question, I simulated another 10,000 samples by bootstrapping the original ratio differences. The bootstrap was made relating the average difference in rev_ratio in relation to the "adwords_pause" variable (treament or control).

**Result**: The histogram shows the average difference in revenue ratio between treatment and controlled groups. 0 is clearly in the confidence interval on both the histogram and in the 95% confidence interval. With 95% confidence, the true average difference between the groups is between `r round(quantile(boot_ebay$diffmean, probs = 0.025),3)` and `r round(quantile(boot_ebay$diffmean, probs = 0.975),3)`.

```{r, echo = FALSE, message = FALSE}
ggplot(boot_ebay) + geom_histogram(aes(x = diffmean), col = "black", bins = 40)
confint(boot_ebay, level = 0.95)
```

**Conclusion**: Because 0 is NOT within the confidence interval of the average difference of revenue_ratio between treatment and controlled groups, there is a statistically significant difference in revenue ratio between the groups. Therefore, the data does in fact favor the idea that paid search advertising on Google creates extra revenue for Ebay.
